A longitudinal analysis of the social information in infants' naturalistic

### 2022-04-21 paper/manuscript/manuscript_april2022_apa_blinded.Rmd
latest revisions to manuscript / figures

We thus capitalize on this computational advance to annotate the entirety of this large, longitudinal video dataset. 
We use these detections to first report how the proportion of faces and hands changes with age in each of the three children in the dataset. 
We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the head cameras as well as a diversity of locations in which videos were recorded.
Finally, making use of automated annotation of pose bounding boxes, we analyze the size, location, and variability of detected faces and poses across development in these three children.
Overall, these findings largely validate prior cross-sectional findings, and further suggest that automated annotations can be used to analyze naturalistic video datasets to build towards generalizable conclusions about the statistics of the social information in the infant view across a wide variety of populations.

@@

We next inspected frames where OpenPose failed to detect faces/hands or erroneously detected them to gain insight into the factors driving performance. Appendix Figure A1 shows randomly sampled examples of both missed detections and false alarms for both faces and hands and highlights the challenging detection conditions presented by these at-home, naturalistic videos. Many missed faces included very far away or very close up views of faces at odd angles, as well as faces that were partially occluded by odd lighting conditions. False alarms for faces included animal faces depicted in picture books and poses where the person was not facing a camera. Hands were sometimes erroneously detected when a pose was present but the hands were out of view, and hands were often missed when they were seen only from the side and unconnected to a pose or belonged to the child. One conclusion of this examination is that egocentric videos using cameras with larger view angles, higher resolution, and more lighting-invariant sensors may lead to better performance, even in the absence of better detection models.

@@

### Accuracy as a function of detection confidence
We next examined the degree to which including detections that the model was more or less confident in changed overall model performance for face/hand detection. When we restricted detections to high-confidence face/hand detections (>.5 confidence, default threshold for visualization in OpenPose) was not beneficial -- improving precision but dramatically impairing recall and thus overall performance: the F-score for high-confidence face detections was `r round(face_performance_hc[3],2)`, with a precision of `r round(face_performance_hc[1],2)` and recall of `r round(face_performance_hc[2],2)`; for high-confidence hand detections, the F-score was `r round(hand_performance_hc[3],2)`, with a precision of `r round(hand_performance_hc[1],2)` and recall of `r round(hand_performance_hc[2],2)`. Further analyses sweeping across the entire range of possible confidence detection thresholds revealed that including all detections was the most beneficial for optimizing the rate of true positives (see Appendix Figure A4 for full ROC curves for both face and hand detections).

### Accuracy across child, location, and age
Finally, we examined whether model accuracy for hands and faces varied across age or between child and the location in which the videos were filmed (e.g., kitchen vs. living room). Some of the videos at taken at older ages for child A did seem to show lower hand detection accuracy, perhaps due the presence of child's hands; however, there were also fewer videos at these time points (see plots for age/child in Appendix, Figure A5). We did not find any discernible trends in face/hand detection accuracy across different filming locations (see Figure A6). However, we did still see considerable variation in model accuracy across different videos, suggesting that particular recording activities, lighting conditions, or situations could cause conditions that would impair detector performance.

@@

```{r faceHandContingency, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of joint face and hands detection within frames where hands (left) or faces (right) were detected." }
ggplot(face_hand_contingency, 
       aes(x=age_day_bin, y=prop, 
           col=child_id)) +
  geom_point(alpha=.1, aes(size=log10(num_frames_total)), show.legend= FALSE) +
  geom_smooth(method='lm', aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion joint hand/face detection') + 
  xlab('Age (Months)') +
  theme_few(base_size=12) +
  facet_grid(~region) +
  # scale_color_manual(values = c( "#DAA0EE","#710D93"), name="") +
  scale_color_brewer(palette="Dark2") +
  guides(size = FALSE) +
  theme(legend.text=element_text(size=10)) 
```

@@

@@

ggplot(avg_area_by_age, aes(x=age_day_bin, y=log(avg_area), col=child_id, size=num_detect)) +
  theme_few(base_size=12) +
  geom_point(alpha=.1) +
  facet_grid(~label) +
  geom_smooth(method='lm') + 
  ylab('Log area (average proportion of frame)') +
  xlab('Age (in months)') +
  scale_color_brewer(palette="Dark2", name = "Child") +
  guides(size = FALSE)
```

@@

@@

@@

cowplot::plot_grid(eye_dist_plot, CV_faces_plot, rel_widths=c(1,2))
```
In a final set of analyses, we explored fine-grained changes in the SAYCam infants' access to social information across development.
In these analyses, we capitalize on the fact that OpenPose provides not only face and hand detections but also positional keypoints. 
In particular, we explored this keypoint dataset with the idea that greater mobility allows older children to be further from their caregivers on average. 
Thus, younger, less mobile children may tend to see larger faces towards the center of their visual field while older, more mobile children may experience more smaller, more variable views of faces.
The same dynamic would be predicted hold for hands as well, as it would be driven by overall differences in distance.

Supporting this idea, we found that the averages sizes of the people, faces, and hands in the infant view became smaller over development (Figure \ref{fig:avgSizes}). 
This effect was relatively consistent across the three children in the dataset, despite the fact that the three children showed sometimes disparate overall proportions of faces/hands in view. 
Thus, children may see closer, larger views of people, hands, and faces earlier in development.

In keeping with this hypothesis, we also found evidence that faces tended to be farther away from older children. 
We restricted our analysis here to faces where both eyes were detected and computed interpupillary distance as a rough metric of distance, since eyes should be closer together on average when a face is further from the camera. 
Figure \ref{fig:faceDetailed}A shows the average interpupillary distance on faces as a function of each child’s age at the time of recording. 
There was a trend from larger, closer faces (with a larger interpupillary distance) to smaller faces that were farther away (with a smaller interpupillary distance).

@@

Finally, we also examined whether there were changes in where faces tended to appear in the camera's (and hence, by proxy, the child's) field of view. 
As expected, faces tended to be located towards the upper field of view, while views of hands were more centrally distributed (see Appendix, Figure A9 for average density distributions). 
However, we also found evidence that older children tended to see more faces in more variable positions than younger children. 
Specifically, we examined how variable the horizontal and vertical coordinates were of the faces in the infant view. 
To do so, we calculated the coefficient of variation of the horizontal (x) and vertical (y) positions of centers of the faces detected by OpenPose (see Figure \ref{fig:faceDetailed}B), and examined changes across age. 
Faces tended to be more variable in the vertical than their horizontal position (see Figure \ref{fig:faceDetailed}B). 
We also found that as children got older, they tended to see faces that varied more in their horizontal -- but not their vertical position -- suggesting that older children might be more likely to see more smaller faces in their periphery (see Figure \ref{fig:faceDetailed}B).

@@

Of course, these automated detections do not entirely obviate the need for careful inspection and manual annotation of dense video data. While the present results suggest that OpenPose achieves moderate performance at detecting faces and hands in infant egocentric videos -- as in @long2021automated -- these results do not guarantee that OpenPose will generalize with similar accuracy to other video datasets.   While OpenPose does provide a considerable amount of about all of the people in a given image -- estimating the entire poses for every individual -- it may not perform accurately enough to answer certain fine-grained research questions. Going forward, we suggest that researchers continue to compare a small set of manual annotations relevant to their research questions of interest to OpenPose's performance.

@@

# Acknowledgements
Blinded.

>>>>

We thus capitalize on this computational advance to annotate the entirety of this large, longitudinal video dataset. 
We use these detections to first report how the proportion of faces and hands changes with age in each of the three children in the dataset. 
We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the head cameras as well as a diversity of locations in which videos were recorded.
Finally, making use of automated annotation of pose bounding boxes, we analyze the size, location, and variability of detected faces and poses across development in these three children.
Overall, these results largely validate prior cross-sectional findings within more naturalistic learning contexts and individual children. These results also highlight some of the challenges of automated coding of infant egocentric headcam data, raising key areas for future computational and empirical work. At the same time, these findings highlight the promise of using automated annotations to analyze large, naturalistic video datasets, advancing the overarching goal of building generalizable conclusions about the statistics of the social information in the infant view across a wide variety of populations.

@@

However, compared to prior work annotating in-lab, egocentric headcam videos, OpenPose performed somewhat less accurately: in Long et al., 2021, the F-score for faces was F = .83 (vs. .64 here), while the F-score for hands was F = .73 (vs. .51 here) (see Table 2, Long et al., 2021). We next inspected frames where OpenPose failed to detect faces/hands or erroneously detected them to gain insight into the factors driving performance. Appendix Figure A1 shows randomly sampled examples of both missed detections and false alarms for both faces and hands and highlights the challenging detection conditions presented by these at-home, naturalistic videos. Many missed faces included very far away or very close up views of faces at odd angles, as well as faces that were partially occluded by odd lighting conditions. False alarms for faces included animal faces depicted in picture books and poses where the person was not facing a camera. Hands were sometimes erroneously detected when a pose was present but the hands were out of view, and hands were often missed when they were seen only from the side and unconnected to a pose or belonged to the child.

Overall, these results suggest that at-home, naturalistic videos taken with these lightweight cameras present unique challenges for applying computer vision methods that may be absent in controlled laboratory settings.  In addition, they suggest that the development and use of head-mounted cameras with larger view angles, higher resolution, and more lighting-invariant sensors may lead to better performance, even in the absence of better detection models.

@@

### Accuracy as a function of detection confidence
We next examined the degree to which including detections that the model was more or less confident in changed overall model performance for face/hand detection. When we restricted detections to high-confidence face/hand detections (>.5 confidence, default threshold for visualization in OpenPose) was not beneficial -- improving precision but dramatically impairing recall and thus overall performance: the F-score for high-confidence face detections was `r round(face_performance_hc[3],2)`, with a precision of `r round(face_performance_hc[1],2)` and recall of `r round(face_performance_hc[2],2)`; for high-confidence hand detections, the F-score was `r round(hand_performance_hc[3],2)`, with a precision of `r round(hand_performance_hc[1],2)` and recall of `r round(hand_performance_hc[2],2)`. This is consistent with findings from Long et al., 2021, where restricting analyses to high-confidence detections also impaired detector performance. Further analyses sweeping across the entire range of possible confidence detection thresholds revealed that including all detections was the most beneficial for optimizing the rate of true positives (see Appendix Figure A4 for full ROC curves for both face and hand detections).

### Accuracy across child, location, and age
Finally, we examined whether model accuracy for hands and faces varied across age or between child and the location in which the videos were filmed (e.g., kitchen vs. living room). Some of the videos at taken at older ages for child A did seem to show lower hand detection accuracy, perhaps due the presence of child's hands; however, there were also fewer videos at these time points (see plots for age/child in Appendix, Figure A5). We did not find any discernible trends in face/hand detection accuracy across different filming locations (see Figure A6). However, we did still see considerable variation in model accuracy across different videos, suggesting that particular recording activities, lighting conditions, or situations could cause conditions that would impair detector performance. For example, challenging lighting conditions in a poorly-lit room or very bright outdoor day could lead to overall lower detection accuracy, as could activities where caregivers are moving rapidly, leading to more blurred views of faces that can be difficult to detect.

@@

```{r faceHandContingency, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of joint face and hands detection within frames where hands (left) or faces (right) were detected." }
ggplot(face_hand_contingency, 
       aes(x=age_day_bin, y=prop, 
           col=child_id)) +
  geom_point(alpha=.1, aes(size=log10(num_frames_total)), show.legend= FALSE) +
  geom_smooth(method='lm', aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion joint hand/face detection') + 
  xlab('Age (Months)') +
  theme_few(base_size=12) +
  ylim(0,1) +
  facet_grid(~region) +
  # scale_color_manual(values = c( "#DAA0EE","#710D93"), name="") +
  scale_color_brewer(palette="Dark2", name = "Child") +
  guides(size = FALSE) +
  theme(legend.text=element_text(size=10))

ggsave('figs_pdfs/face_hand_contingency.pdf', width=7, height=3, units='in')
```

@@

ggsave('figs_pdfs/dets_by_location.pdf', width=7, height=5, units='in')
```

@@

ggplot(avg_area_by_age, aes(x=age_day_bin, y=log(avg_area), col=child_id, size=num_detect)) +
  theme_few(base_size=12) +
  geom_point(alpha=.1) +
  facet_grid(~label) +
  geom_smooth(method='lm') + 
  ylab('Log area (average proportion of frame)') +
  xlab('Age (in months)') +
  scale_color_brewer(palette="Dark2", name = "Child") +
  guides(size = FALSE)

ggsave('figs_pdfs/size_by_age.pdf', width=7, height=3, units='in')
```

```{r}
body_size_summary = summary(lmer(data = all_vid_bbs %>% filter(label=='pose'), log(avg_area) ~ scale(age_days) + scale(num_detect) + (1|child_id)))

faces_size_summary = summary(lmer(data = all_vid_bbs %>% filter(label=='face'), log(avg_area) ~ scale(age_days) + scale(num_detect) + (1|child_id)))

hands_size_summary = summary(lmer(data = all_vid_bbs %>% filter(label=='hand'), log(avg_area) ~ scale(age_days) + scale(num_detect) + (1|child_id)))

@@

```{r}
eye_dist_summary = summary(lmer(data = all_vid_data, avg_eye_distance ~ scale(age_days) + scale(num_frames) + (1|child_id)))
```

@@

```{r}
x_vs_y_summary = summary(lmer(data=CV_faces, CV ~ scale(age_day_bin) + position + scale(num_detect) +  (1|child_id)))

@@

cowplot::plot_grid(eye_dist_plot, CV_faces_plot, rel_widths=c(1,2))
ggsave('figs_pdfs/detailed_changes.pdf',units='in',width=7, height=3)
```
In a final set of analyses, we explored fine-grained changes in the SAYCam infants' access to social information across development.
In these analyses, we capitalize on the fact that OpenPose provides not only face and hand detections but also positional keypoints. 
In particular, we explored this keypoint dataset with the idea that greater mobility allows older children to be further from their caregivers on average. 
Thus, younger, less mobile children may tend to see larger faces towards the center of their visual field while older, more mobile children may experience more smaller, more variable views of faces.
The same dynamic would be predicted hold for hands as well, as it would be driven by overall differences in distance.

Supporting this idea, we found that the averages sizes of the people, faces, and hands in the infant view became smaller over development (Figure \ref{fig:avgSizes}). To assess these trends, we fit three linear mixed-effect models, asking whether the average sizes of people, faces, and hands declined with children’s age, and found significant effects in each case (People, B = `r round(body_size_summary$coefficients,4)[2,1]`,
SE = `r round(body_size_summary$coefficients,4)[2,2]`, df = `r round(body_size_summary$coefficients,4)[2,3]`, t = `r round(body_size_summary$coefficients,4)[2,4]`, 
p = `r round(body_size_summary$coefficients,4)[2,5]`;Faces, B = `r round(faces_size_summary$coefficients,3)[2,1]`,
SE = `r round(faces_size_summary$coefficients,4)[2,2]`, df = `r round(faces_size_summary$coefficients,4)[2,3]`, t = `r round(faces_size_summary$coefficients,4)[2,4]`,
p = `r round(faces_size_summary$coefficients,4)[2,5]`); Hands, B = `r round(hands_size_summary$coefficients,4)[2,1]`,
SE = `r round(hands_size_summary$coefficients,4)[2,2]`, df = `r round(hands_size_summary$coefficients,4)[2,3]`, t =`r round(hands_size_summary$coefficients,4)[2,4]`, p = `r round(hands_size_summary$coefficients,4)[2,5]`).

This effect was relatively consistent across the three children in the dataset, despite the fact that the three children showed sometimes disparate overall proportions of faces/hands in view. 
Thus, children may see closer, larger views of people, hands, and faces earlier in development.

In keeping with this hypothesis, we also found evidence that faces tended to be farther away from older children. 
We restricted our analysis here to faces where both eyes were detected and computed interpupillary distance as a rough metric of distance, since eyes should be closer together on average when a face is further from the camera. 
Figure \ref{fig:faceDetailed}A shows the average interpupillary distance on faces as a function of each child’s age at the time of recording. 
There was a trend from larger, closer faces (with a larger interpupillary distance) to smaller faces that were farther away (with a smaller interpupillary distance);  this finding was confirmed in a linear-mixed effect model (B = `r round(eye_dist_summary$coefficients,4)[2,1]`, SE = `r round(eye_dist_summary$coefficients,4)[2,2]`, df = `r round(eye_dist_summary$coefficients,4)[2,3]`, t = `r round(eye_dist_summary$coefficients,4)[2,4]`, P = `r round(eye_dist_summary$coefficients,4)[2,5]`).

@@

Finally, we also examined whether there were changes in where faces tended to appear in the camera's (and hence, by proxy, the child's) field of view. 
As expected, faces tended to be located towards the upper field of view, while views of hands were more centrally distributed (see Appendix, Figure A9 for average density distributions). 
However, we also found evidence that older children tended to see more faces in more variable positions than younger children. 
Specifically, we examined how variable the horizontal and vertical coordinates were of the faces in the infant view. 
To do so, we calculated the coefficient of variation of the horizontal (x) and vertical (y) positions of centers of the faces detected by OpenPose (see Figure \ref{fig:faceDetailed}B), and examined changes across age. 
Faces tended to be more variable in the vertical than their horizontal position (see Figure \ref{fig:faceDetailed}B, Main effect of position type on variability, B = `r round(x_vs_y_summary$coefficients,4)[3,1]`, SE = `r round(x_vs_y_summary$coefficients,4)[3,2]`, df = `r round(x_vs_y_summary$coefficients,4)[3,3]`, t = `r round(x_vs_y_summary$coefficients,4)[3,4]`, P = `r round(x_vs_y_summary$coefficients,4)[3,5]`).
We also found that as children got older, they tended to see faces that varied more in their horizontal -- but not their vertical position -- suggesting that older children might be more likely to see more smaller faces in their periphery (see Figure \ref{fig:faceDetailed}B), (Main effect of age of horizontal position, B = `r round(x_pos_summary$coefficients,4)[2,1]`, SE = `r round(x_pos_summary$coefficients,4)[2,2]`, df = `r round(x_pos_summary$coefficients,4)[2,3]`, t = `r round(x_pos_summary$coefficients,4)[2,4]`, P = `r round(x_pos_summary$coefficients,4)[2,5]`; No main effect of age on vertical position, Main effect of age of horizontal position, B = `r round(y_pos_summary$coefficients,4)[2,1]`, SE = `r round(y_pos_summary$coefficients,4)[2,2]`, df = `r round(y_pos_summary$coefficients,4)[2,3]`, t = `r round(y_pos_summary$coefficients,4)[2,4]`, P = `r round(y_pos_summary$coefficients,4)[2,5]`; )

@@

Of course, these automated detections do not entirely obviate the need for careful inspection and manual annotation of dense video data. While the present results suggest that OpenPose achieves moderate performance at detecting faces and hands in infant egocentric videos -- as in @long2021automated -- these results do not guarantee that OpenPose will generalize with similar accuracy to other video datasets.   While OpenPose does provide a considerable amount of about all of the people in a given image -- estimating the entire poses for every individual -- it may not perform accurately enough to answer certain fine-grained research questions. Going forward, we suggest that researchers continue to compare a small set of manual annotations relevant to their research questions of interest to OpenPose's performance. Indeed, different cameras, lighting conditions, or recording environments may change the accuracy of the detections and thus the inferences that can be made. In particular, we suggest that researchers continue to annotate a set of randomly sampled frames from each child across an entire dataset to assess the rates of both misses as well as false alarms. While we recommend at least a few thousand frames be annotated, additional annotations will allow for more precise estimates of model performance in a given dataset.

@@

# Acknowledgements
Thanks to the creators of the SAYCam dataset who made this work possible and to Alessandro Sanchez for his contributions to the codebase. This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL.
